---
title: "AMES Housing Data - EDA"
output:
  html_document:
    df_print: paged
---
Import packages and the dataset.
```{r}
library(ggplot2)
library(epiDisplay)
library(dplyr)
library(visdat)
library(corrplot)
options(scipen=10000)
```
Import the dataset.
```{r}
df <- data.frame(AMES_TRAIN)
head(df)
```


# Define the sample

1. Select the variables of interest.

```{r}
df <- df %>% 
  select(Zoning, LotFrontage, LotArea, BldgType, OverallQual, OverallCond, YearBuilt, YearRemodel, ExterQual, FirstFlrSF, SecondFlrSF, FullBath, HalfBath, BsmtFullBath, BsmtHalfBath, BedroomAbvGr, KitchenQual, GarageArea, SaleCondition, SalePrice, index)

df
```
2. let's look at how many records are missing values in our dataset.

```{r}
vis_miss(df)
```

3. Subset only properties that are Residential

```{r}
df <- subset(df, df$Zoning == 'RH' | df$Zoning == 'RL' | df$Zoning == 'RP' | df$Zoning == 'RM')
```

4. Subset only properties that had normal sale conditions.

```{r}
df <- subset(df, df$SaleCondition == 'Normal')
```


5. Examine BldgType with a frequency table and histogram. 
```{r}
tab1(df$BldgType, sort.group = "decreasing", cum.percent = TRUE)
```
__Keep only records that are BldgType = 1Fam.__

```{r}
df <- df %>% filter(BldgType == "1Fam" | is.na(BldgType))
```

6. Examine SalePrice and remove outliers. 

```{r}
summary(df$SalePrice)
ggplot(df, aes(x=SalePrice)) + geom_histogram(fill='sky blue')
ggplot(df, aes(y=SalePrice)) + geom_boxplot(fill='sky blue')
```
ggplot2 boxplots define outliers above the median as values greater than Q3 + 1.5*IQR. [source - stackoverflow](https://stackoverflow.com/questions/4946964/in-ggplot2-what-do-the-end-of-the-boxplot-lines-represent)

Let's use the [boxplot.stats](https://stat.ethz.ch/R-manual/R-patched/library/grDevices/html/boxplot.stats.html) function to find the minimum, Q2, Median, Q3, and max of the upper whisker. __boxplot.stats__ calculates whisker extremes as: +/-1.58 IQR/sqrt(n) 

```{r}
bp_stats <- boxplot.stats(df$SalePrice)[1]
bp_labels <- c('extreme lower whisker','Q2','median','Q3','extreme upper whisker')
bp_df <- data.frame(bp_stats,bp_labels)
bp_df
```
Now that we have SalePrice outliers defined as above the extreme upper whisker at 345474, lets filter out records above this extreme upper whisker.

```{r}
df <- df %>% filter(SalePrice <= 345474 | is.na(SalePrice))
```


8. Calculate total square foot area

```{r}
df$TotSF <- df$FirstFlrSF + df$SecondFlrSF
```


9. Calculate TotalBath
```{r}
# drop NAs
df <- df[!with(df, is.na(BsmtFullBath) & is.na(BsmtHalfBath)),]
df$TotalBath <- df$FullBath + df$HalfBath + df$BsmtFullBath + df$BsmtHalfBath
```

10. Check to see if any SalePrice values below zero.

```{r}
df[df$SalePrice < 0,]
```


11. LotFrontage is the only variable with considerable missing values (18.81% missing).

Let's compare LotFrontage to SalePrice to see if this appears to be a good variable to use in our final regression model.

```{r}
ggplot(df, aes(x=LotFrontage,y=SalePrice)) + 
  geom_point(shape=16, color='royalblue3') +
  geom_smooth(method=lm,  linetype="dashed", color="darkred", fill="blue")
```
LotFrontage looks promising. 

Let's compare LotFrontage to LotArea.

```{r}
ggplot(df, aes(x=LotArea, y=LotFrontage)) + 
  geom_point(shape=16, color='royalblue3') +
  geom_smooth(method=lm,  linetype="dashed", color="darkred", fill="blue")
```
Let's fill in the NA values in LotFrontage with predicted values from LotArea using a regression model. 

```{r}
# copy the dataframe so that we are not working on the original
lf_imp <- df %>% 
  select(LotFrontage, LotArea) %>% # select rows
  na.omit() # drop NAs

# Create a linear regression model predicting LotFrontage from LotArea
lm(LotFrontage ~ LotArea, lf_imp)
```
We got an intercept of 48.604608 and slope of 0.002338.

In our original dataset df, let's create a dummy variable for if LotFrontage is NA, using [this code from Rpub.](https://rpubs.com/GourabNath/MissingValue), and predict LotFrontage from LotArea if the value is NA.

```{r}
# Create a function that will label our dummy variables for a given column
label_dummies <- function(t)
{
  x <- dim(length(t)) 
  x[which(!is.na(t))] = 1
  x[which(is.na(t))] = 0
  return(x)
}

# Call our function on the dataframe
df$LFdummy <- label_dummies(df$LotFrontage)

# Impute the missing values using our model.

for(i in 1:nrow(df))
{
  if(df$LFdummy[i] == 0)
  {
    df$LotFrontage[i] = 48.604608 + 0.002338*df$LotArea[i]
  }
}
```

While we are working on LotFrontage, let's also drop the one record where LotFrontage is greater than 300 feet, as it has an extremely low SalePrice.

```{r}
df <- subset(df, df$LotFrontage < 300)
```

# Exploratory Data Analysis - Continuous Variables
Subset all the continuous variables.

```{r}
df_con <- df %>% select(LotArea, YearBuilt, TotSF, LotFrontage, GarageArea, SalePrice)
```

Create a correlogram of the continuous variables.
```{r}
library(corrplot)

corrplot(cor(df_con), type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, addCoef.col = "white")
```
Let's investigate the three highest correlations to SalePrice: YearBuilt, GarageArea, TotSF, and LotArea.

X = YearBuilt
```{r}
ggplot(df_con, aes(x=YearBuilt, y=SalePrice)) + 
  geom_point(shape=16, color='royalblue3') +
  geom_smooth(method=lm,  linetype="dashed", color="darkred", fill="blue")
```
X = GarageArea
```{r}
ggplot(df_con, aes(x=GarageArea, y=SalePrice)) + 
  geom_point(shape=16, color='royalblue3') +
  geom_smooth(method=lm,  linetype="dashed", color="darkred", fill="blue")
```
X = TotSF

```{r}
ggplot(df_con, aes(x=TotSF, y=SalePrice)) + 
  geom_point(shape=16, color='royalblue3') +
  geom_smooth(method=lm,  linetype="dashed", color="darkred", fill="blue")
```
X = LotArea
```{r}
ggplot(df_con, aes(x=LotArea, y=SalePrice)) + 
  geom_point(shape=16, color='royalblue3') +
  geom_smooth(method=lm,  linetype="dashed", color="darkred", fill="blue")
```
Let's remove outliers from LotArea, and redraw scatterplot
```{r}
df_con_LA <- df_con[df_con$LotArea < 30000,]
ggplot(df_con_LA, aes(x=LotArea, y=SalePrice)) + 
  geom_point(shape=16, color='royalblue3') +
  geom_smooth(method=lm,  linetype="dashed", color="darkred", fill="blue")
cor(df_con_LA$LotArea, df_con_LA$SalePrice)
```

# Exploratory Data Analysis - Discrete Variables

```{r}
df_disc <- df %>% select(TotalBath, BedroomAbvGr, SalePrice)
```

Boxplot of TotalBath
```{r}
p <- df_disc %>% mutate(TotalBath = as.character(TotalBath))
p<-ggplot(p, aes(x=TotalBath, y=SalePrice, color=TotalBath)) +
  geom_boxplot()
p
```
Boxplot of BedroomAbvGr

```{r}
p <- df_disc %>% mutate(BedroomAbvGr = as.character(BedroomAbvGr))
p<-ggplot(p, aes(x=BedroomAbvGr, y=SalePrice, color=BedroomAbvGr)) +
  geom_boxplot()
p
```
```{r}
df[df$BedroomAbvGr == 0,]
```
# Exploratory Data Analysis - Ordinal Variable
```{r}
df_ord <- df %>% select(OverallQual, ExterQual, KitchenQual, SalePrice)
```

Boxplot of OverallQual
```{r}
p <- df_ord %>% mutate(OverallQual = as.character(OverallQual))
p<-ggplot(p, aes(x=OverallQual, y=SalePrice, color=OverallQual)) +
  geom_boxplot()
p
```
Boxplot of ExterQual

```{r}
p <- df_ord %>% mutate(ExterQual = as.character(ExterQual))
p<-ggplot(p, aes(x=ExterQual, y=SalePrice, color=ExterQual)) +
  geom_boxplot()
p
```
Boxplot of KitchenQual
```{r}
p <- df_ord %>% mutate(KitchenQual = as.character(KitchenQual))
p<-ggplot(p, aes(x=KitchenQual, y=SalePrice, color=KitchenQual)) +
  geom_boxplot()
p
```
# Models

Total Square Feet - TotSF

```{r}
totsf_lm <- lm(SalePrice ~ TotSF, data=df)
totsf_lm 
summary(totsf_lm)
```
TotalBath

```{r}
totalbath_lm <- lm(SalePrice ~ TotalBath, data=df)
totalbath_lm
summary(totalbath_lm)
```
# Multiple regression

TotSF and TotalBath

```{r}
sf_bath_lm <- lm(SalePrice ~ TotSF+TotalBath, data=df)
sf_bath_lm
summary(sf_bath_lm)
```
TotSF and LotFrontage
```{r}
sf_lf_lm <- lm(SalePrice ~ TotSF+LotFrontage, data=df)
sf_lf_lm
summary(sf_lf_lm)
```

# Predict SalePrice

Read in test data set.

```{r}
test <- data.frame(AMES_TEST_SFAM)

# create the calculated variables TotSF and TotalBath
test$TotSF <- test$FirstFlrSF + test$SecondFlrSF
test$TotalBath <- test$FullBath + test$HalfBath + test$BsmtFullBath + test$BsmtHalfBath
```

Predict values

```{r}
df1 <- test[ ,c("index","p_saleprice","SalePrice","TotSF","TotalBath")]
p_saleprice <- predict(sf_bath_lm, df1)
df_out <- data.frame(p_saleprice,test$index)
head(df_out, n=5)
```

```{r}
write.csv(df_out,'~/OneDrive/NORTHWESTERN/Machine Learning/Module 1 - EDA/homework/predicted_test.csv',row.names=TRUE)
```





